Discovery of the first quadruple gravitationsally lensed quasar candidate with Pan-STARRS

### 2017-05-11 paper4.tex
Add files via upload
First commit

\author{C. T. Berghea\altaffilmark{1}, G. Nelson\altaffilmark{1}, C. E. Rusu\altaffilmark{2}, C. R. Keeton\altaffilmark{3}}

@@

We report the serendipitous discovery of the first quadruple gravitational lens candidate in the Pan-STARRS images. The lensing galaxy is detected by fitting the lens in i, z and y bands using the Hostlens software. There are small differences in the colors of the four quasars images but they are generally consistent with a low redshift quasar at about z $\sim$ 0.8 based on photometric redshift estimates. We model the lens with both an isothermal ellipsoid (SIE) or a isothermal sphere with shear. While both models provide relatively good fits to the data, there are important discrepancies in the fluxes, especially betwen the B and C components which the models predict should show similar fluxes. These can be explained by differential extinction, microlensing by stars and millilensing due to the dark matter substructure. 
\end{abstract}

@@

To calculate photometric redshift we used the method of \citet{wu10}, which is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citep{fin16} to obtain magnitudes in the SDSS system. Folowing \citet{wu10} we minimize ${\chi}^2$ to obtain redshits for each of the sources. We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image and also fitting all togehter. In the latter case we obtain the best fit at z $\sim$ 0.83, but this is not well constrained, the 1~$\sigma$ confidence level is (0.625,  2.875). We note that there is a second minimum at z $\sim$ 2.6. $\chi^2_{\nu}$ is 0.56 and 1.07 for the two minima, respectively.

@@

Due to blending of A, B and C components, both Gaia \citep{gaia} and PS1 catalogs identified sources A and D only. 
PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023. Gaia G magnitudes are 15.17 and 18.13 for A and D, respectively.
We will use the Gaia position for component D as absolute astrometric reference position for our lens: 26.7923069811, 46.5112731937. The errors for this position are 15.6 and 8.9 mas, respectively.

@@

g & 15.60$\pm$0.01 &  15.72$\pm$0.01 &  16.45$\pm$0.02 & 18.09$\pm$0.01 & $\cdots$ \\
r & 15.40$\pm$0.01 &  15.55$\pm$0.01 &  16.21$\pm$0.01 & 17.74$\pm$0.01 & $\cdots$ \\
i & 15.36$\pm$0.01 &  15.57$\pm$0.02 &  16.15$\pm$0.02 & 17.74$\pm$0.02 & 19.50$\pm$0.20\\
z & 15.23$\pm$0.03 &  15.50$\pm$0.05 &  16.02$\pm$0.01 & 17.68$\pm$0.03 & 18.95$\pm$0.13\\
y & 14.92$\pm$0.01 &  15.23$\pm$0.02 & 15.76$\pm$0.02 & 17.36$\pm$0.02 & 19.20$\pm$0.24\\
$\Delta$i & 0.0 & 0.21 & 0.79 & 2.38 & 4.14 \\
$\Delta$x & 0.000$\pm$0.004 & -1.185$\pm$0.004 & 1.271$\pm$0.005 & 0.410$\pm$0.004 & 0.240$\pm$0.050\\
$\Delta$y & 0.000$\pm$0.004 & -0.441$\pm$0.004 & -0.074$\pm$0.004 & -3.310$\pm$ 0.004
& -2.310$\pm$0.025\\
z & 0.725 & 0.925 & 0.925 & 1.125 \\

@@

\tablecomments{
The first four columns are the four quasar images and the fifth column is the lens.
The first five rows are magnitude measurements derived using Galfit and recalibrated using PS1 photometry for nearby stars and corrected using NED extinction calculator.
The following three rows are relative i magnitudes and relative positions in arcseconds obtained from Galfit. The reference magnitude is the i magnitude of source A and reference position is the average between the i and y bands for source A.
The next row shows the photometric redshift estimates following \citet{wu10}.
Following rows shows the model prediction for position and relative magnitudes for the two gravitational models we tried.
}
\end{deluxetable}

@@

\begin{figure}
\epsscale{1.1}
\plottwo{redshift5.eps}{redshift4.eps}
\caption{
Photometric redshift estimates. 
Left: individually for each image; right: for all images simultanesously}
\label{redshift}
\end{figure}

>>>>

\author{C. T. Berghea\altaffilmark{1}, G. Nelson\altaffilmark{1}, C. E. Rusu\altaffilmark{2}, C. R. Keeton\altaffilmark{3}, R. P. Dudik\altaffilmark{1}}

@@

We report the serendipitous discovery of the first quadruple gravitational lens candidate in the Pan-STARRS images. The lensing galaxy is detected by fitting the lens in i, z and y bands using the Hostlens software. There are small differences in the colors of the four quasars images but they are generally consistent quasar at about z $\sim$ 2.6 based on photometric redshift estimates. The lensing galaxy is very faint but we used galaxy templates to estimate z $\sim$ 0.6. We model the lens with both an isothermal ellipsoid (SIE) or a isothermal sphere with shear. While both models provide relatively good fits to the data, there are important discrepancies in the fluxes, especially betwen the B and C components which the models predict should show similar fluxes. These can be explained by differential extinction, microlensing by stars and millilensing due to the dark matter substructure. 
\end{abstract}

@@

{\bf This has to be changed to use Le Phare method }
To calculate photometric redshift we used the method of \citet{wu10}, which is based on Sloan Digital Sky survey (SDSS) colors. PS1 bands are similar but not identical to the SDSS bands and we used the corrections of \citep{fin16} to obtain magnitudes in the SDSS system. Folowing \citet{wu10} we minimize ${\chi}^2$ to obtain redshits for each of the sources. We plot $\chi^2_{\nu}$ in Figure~\ref{redshift} for each quasar image and also fitting all togehter. In the latter case we obtain the best fit at z $\sim$ 0.83, but this is not well constrained, the 1~$\sigma$ confidence level is (0.625,  2.875). We note that there is a second minimum at z $\sim$ 2.6. $\chi^2_{\nu}$ is 0.56 and 1.07 for the two minima, respectively.

@@

Due to blending of A, B and C components, both Gaia \citep{gaia} and PS1 catalogs identified sources A and D only. 
PS1 psf magnitudes for A are in general smaller than what we obtained probably due to blending, but for D they are very similar to our uncorrected magnitudes: 18.599$\pm$0.013,  18.153$\pm$0.002,  18.079$\pm$0.007,  17.843$\pm$0.0093,  17.534$\pm$0.023. Gaia G magnitudes are 15.17 and 18.13 for A and D, respectively.
We will use the Gaia position for component D as absolute astrometric reference position for our lens: 26.7923069811, 46.5112731937. The errors for this position are 15.6 and 8.9 mas, respectively.

@@

g & 15.60$\pm$0.01 &  15.72$\pm$0.01 &  16.45$\pm$0.02 & 18.09$\pm$0.01 & $\cdots$ \\
r & 15.40$\pm$0.01 &  15.55$\pm$0.01 &  16.21$\pm$0.01 & 17.74$\pm$0.01 & $\cdots$ \\
i & 15.36$\pm$0.01 &  15.57$\pm$0.02 &  16.15$\pm$0.02 & 17.74$\pm$0.02 & 19.50$\pm$0.20 \\
z & 15.23$\pm$0.03 &  15.50$\pm$0.05 &  16.02$\pm$0.01 & 17.68$\pm$0.03 & 18.95$\pm$0.13 \\
y & 14.92$\pm$0.01 &  15.23$\pm$0.02 & 15.76$\pm$0.02 & 17.36$\pm$0.02 & 19.20$\pm$0.24 \\
$\Delta$i & 0.0 & 0.21 & 0.79 & 2.38 & 4.14 \\
$\Delta$x & 0.000$\pm$0.004 & -1.185$\pm$0.004 & 1.271$\pm$0.005 & 0.410$\pm$0.004 & 0.240$\pm$0.050 \\
$\Delta$y & 0.000$\pm$0.004 & -0.441$\pm$0.004 & -0.074$\pm$0.004 & -3.310$\pm$ 0.004 & -2.310$\pm$0.025 \\
z & 0.725 & 0.825 & 0.725 & 0.925 \\

@@

\tablecomments{
The first four columns are the four quasar images and the fifth column is the lens.
The first five rows are magnitude measurements derived using Hostlens and corrected using NED extinction calculator.
The following three rows are relative i magnitudes and relative positions in arcseconds averaged between the i and y bands. 
The next row shows the photometric redshift estimates following \citet{wu10}.
Following rows shows the model prediction for position and relative magnitudes for two gravitational models.
}
\end{deluxetable}

@@

\begin{figure}
\epsscale{1.1}
\plottwo{redshiftseparate.eps}{redshifttogether.eps}
\caption{
Photometric redshift estimates. 
Left: individually for each image; right: for all images simultanesously}
\label{redshift}
\end{figure}
